Faithful Variable Screening for\\ High-Dimensional Convex Regression

### 2015-05-04 tex/scam/reviews-v1/response-v1.tex

The key ideas behind our analysis do not depend critically on the fact that the support is a hypercube. The hypercube assumption is made for convenience, to minimize technical details.

\item \rc{Intuitively, variable selection is very closely related to
estimating a function (or possibly its partial derivatives) in the
sup-norm.  It is well-known that accurate estimation of functions
(whether in $L^2$ or the sup-norm) with bounded smoothness in dimension
$s$ requires a sample size that grows exponentially in $s$.  For this
reason, the past result of Comminges and Dalalyan (suggesting an
exponential growth for variable selection) is to be expected.
However, in terms of metric entropy, the class of convex functions
behaves roughly like the class of two-smooth functions (e.g., Dryanov
(2009)).  So at first glance, this makes the result of this paper
rather surprising---namely, that one can somehow side-step any
exponential growth in $s$, even though the metric entropy is still
scaling as badly as the case of 2-smooth functions.  I wondered if the
authors could comment and shed insight on this fact, which seems a
little counterintuitive.  What I suspect that it could mean is that
the additive faithfulness (AF) condition, or one of the sufficient
conditions for AF (namely boundary flatness) are somehow very
restrictive.  Could the authors comment on these issues?}

The boundary flatness doesn't artificially make the variable selection easy. Boundary flatness is much more general than the product density and Comminges and Dalalyan showed that, even under a product density, variable selection for general two-smooth functions is hard. Any bounded density with compact support can be approximated arbitrary well by a boundary flat density (see the newly added Example 3.2).

The intuition behind our result is the following: start with the fact that convex functions are additively faithful under a product density. The behavior of a convex function everywhere is constrained by its behavior at the boundary. Therefore, we need only that the underlying density resembles a product density at the boundary, which is exactly the notion that boundary flatness formalizes.

@@

If the true regression function is a convex additive function, then additive faithfulness holds trivially--the best additive approximation to an additive function is the function itself. The estimation procedure becomes simpler--the DC stage is no longer necessary and can be safely removed.

In the finite sample analysis, the assumptions would not change except that A3 would be $\| f^*_k \|_\infty \leq B$ for all $k$ since $f_0 = f^*$ in this case. The signal strength definition can be made simpler: $\alpha^+ = \min_k \mathbb{E}( f^*_k(X_k)^2 )$. The rates would not improve.

\item \rc{If we actually wish to fit smooth convex univariate components, is
it easy to add this condition as a regularization term and incorporate
it into the analysis?}

Smoothness can be incorporated by using B-spline basis with uniformly spaced knots. The details are described in a 2014 paper by Pya and Wood which we have cited in the introduction. We chose not to add smoothness into the estimation because it would introduce an additional tuning parameter.

The analysis may require much more work if one uses a B-spline basis in the estimation. B-splines may complicate the false positive analysis, which rely on the particular form of the optimization objective. The false negative analysis may not be affected much because it mainly uses bracketing entropy.

@@

%\item \rc{On page 10, on line 16, ``additively'' should be replaced by
%``additive''.}

%\item \rc{On page 11, at the bottom of the page, ``before we presenting''}

\item \rc{On page 11, after Deônition 3.2, it is written ``when the joint
density satisfies the property that
$\frac{\partial(x_{-j},x_j)}{\partial x_j} = \frac{\partial^2 p(x_{-j},
  x_j)}{\partial x_j^2} = 0$ at boundary points''.
I think the condition ``$p(x_{−j}, x_j ) = 0$ at the boundary points'' is
missing.}

It is necessary that $p(\mathbf{x}_{-j}, x_j) > 0$. We have added an additional condition that $p(\mathbf{x})$ has a bounded second derivative. This condition ensures that the marginal density $p(x_j)$ is twice differentiable and $p'(x_j) = p''(x_j) = 0$ at $x_j = 0,1$. With this, it is valid to apply the quotient rule for differentiation to show that boundary flatness holds. Please see Example 3.2 for the relevant edits.

@@

%\item \rc{On page 26, in the first paragraph ``the relevant variable set'' and
%in the third paragraph, ``seting''.}

\item \rc{I really did not understand what was represented in figures 5(a) and
5(b) (axis?). Besides it does not seem to correspond with what is
described at the bottom of page 26 and at the top of page 27. The
notation $\|\cdot\|_{\infty,1}$ is not defined.}

We have further explained the old figure 5a and 5b (now figure 7a and 7b) in the experiment section. The following has been added to the paper:

``... we plot on the y-axis the norm $\|f^{(t)}_j\|_{\infty}$ of every column $j$ against the regularization strength $\lambda^{(t)}$. Instead of plotting the value of $\lambda^{(t)}$ on the x-axis however, we plot the total norm at $\lambda^{(t)}$ normalized against the total norm at $\lambda^{(1)}$: $\frac{\sum_j \|f^{(t)}_j\|_{\infty}}{\sum_j \|f^{(1)}\|_{\infty}}$. Thus, as x-axis moves from 0 to 1, the regularization goes from strong to weak.''

@@

\subsection*{Reviewer 3}

Major comments:

\begin{enumerate}
\item \rc{The basis of all the results seem to be the boundary flatness
condition. This condition seems to be an artifact of the analysis and
based on my reading, it is unclear how restrctive this assumption
is. The authors claim it is weak just after stating it and then
present several straightforward examples in which it doesn't hold
(e.g. $f(x_1,x_2)=x_1 x_2$) and multivariate Gaussian examples. Hence it is unclear to me how
likely a given function is likely to fit into this framework.}

The boundary flatness condition is not restrictive in the sense that for any density $p(x)$ supported on $[0,1]^p$, and for any $\epsilon > 0$, there exists a boundary flat density whose $L_1$ distance with $p(x)$ is at most $\epsilon$ (Please see the newly added Example 3.3). This is because boundary flatness places no restriction on what $p(x)$ looks like in the interior of the hypercube. Boundary flatness allows arbitrary correlation structure, provided that $p(x) > 0$. We have added Section 3.2 to the paper to give a more detailed discussion on boundary flatness.

@@

$f(x_1,x_2) = x_1 x_2$ is non-convex and it is an example of how non-convex functions may not be additively faithful under the uniform distribution; it is not an example in which boundary flatness doesn't hold. The multivariate Gaussian is a counterexample but it is to show that boundary flatness may not extend easily to densities with unbounded support.

We acknowledge that many distributions are not boundary flat. We present boundary flatness as a significant generalization of the product density, under which nonparametric variable selection is already very difficult.

Boundary flatness is not a contrived condition. We initially proved additive faithfulness only for product densities. Then, knowing that the behavior of a convex function everywhere is constrained by its behavior at the boundary, we conjectured that underlying density may only need to ``look like'' a product density at the boundary. Attempts to formalize this notion led us to the definition of boundary flatness.

### 2015-05-05 tex/scam/reviews-v1/response-v1.tex

\subsection*{Associate Editor}

\rc{Below I recapitulate a few
point of the reviewers', along with some questions/comments of my own.
These (as well as other points in the reviews) need to be thoroughly
addressed in the revised manuscript.}

\begin{enumerate}
\item \rc{Boundary flatness condition: much of the paper's analysis hinges on
this condition.  Reviewer 3 raises some concerns about when it holds,
especially for functions that are not necessarily supported on the
unit hypercube.  See also the point below.}

We have added a new Section 3.2 to the paper to further discuss the boundary flatness condition as well as giving more examples. Extending our analysis to densities who support is not hypercube is beyond the scope of our paper although we think it is do-able. The key ideas behind our analysis do not depend critically on the fact that the support is a hypercube. The hypercube assumption is made for convenience, to minimize technical details.

### 2015-05-15 tex/scam/additive_faithful.tex
minor

The boundary flatness condition intuitively states that two conditional densities $p(\mathbf{x}_{-j} \given x_j)$ and $p(\mathbf{x}_{-j} \given x_j')$ are similar when $x_j$ and $x_j'$ are both close to the same boundary point. It is thus much more general than product densities. Boundary flatness is a weak condition because it affects only an $\epsilon$-small region around the boundary; $p(\mathbf{x}_{-j} \given x_j)$ can take arbitrary shapes away from the boundary. Boundary flatness also allows arbitrary correlation structure between the variables (provided $p(\mathbf{x}) >0$). In Section~\ref{sec:boundary_flat}, we give a detailed discussion of the boundary flatness condition and show examples of boundary flat densities; in particular, we show that any density supported on a compact set can be approximated arbitrarily well by boundary flat densities.

@@

\begin{remark}
In Theorem~\ref{thm:convex_faithful}, we do not assume a parametric form for the additive components; the additive approximations may not be faithful if we take a parametric form. For example, suppose we approximate a mean-zero convex function $f(X)$ by a linear form $X \beta$. The optimal linear function in the population setting is $\beta^* = \Sigma^{-1} \trm{Cov}(X, f(X))$ where $\Sigma$ is the covariance matrix. Suppose the $X$'s are independent, follow a symmetric distribution, have unit variance, and suppose $f(\mathbf{x}) = x_1^2 - \E[X_1^2]$, then $\beta^*_1 = \E[ X_1 f(X)] = \E[ X_1^3 - X_1 \E[X_1^2]] = 0$.
\end{remark}

@@

In this section, we give more examples of boundary flat densities (see Definition~\ref{defn:boundary-point}) and discuss extending the notion of boundary flatness to densities with a more general support. We first start with an sufficient condition on the \emph{joint density} that ensures boundary flatness.

@@

In our discussion so far, we have restricted ourselves to densities supported and positive on the hypercube $[0,1]^p$ to minimize extraneous technical details. It may be possible to extend the analysis to densities whose support is a convex and compact set so long as the marginal density $p(x_j) > 0$ for all $x_j$ in the support. A rigorous analysis of this however is beyond the scope of this paper.

It may also possible to extend similar result to densities with an
unbounded support, by using a limit condition $\lim_{|x_k| \rightarrow
  \infty} \frac{\partial p(\mathbf{\scriptstyle x}_{-k} \given x_k)}{\partial x_k}
= 0$.  Such a limit condition however is not obeyed by a correlated multivariate Gaussian distribution. 
The next example shows that certain convex functions are not additively faithful under certain multivariate Gaussian distributions.

\begin{example}
\label{examp:gaussian_counterexample}
Consider a two dimensional quadratic function $f( \mathbf{x}) = \mathbf{x}^\tran H \mathbf{x} + c$ with zero mean where $H = \begin{pmatrix} H_{11} & H_{12} \\ H_{12} & H_{22}\end{pmatrix}$ is positive definite and a Gaussian distribution $X \sim N(0, \Sigma)$ where $\Sigma = \begin{pmatrix}1 & \alpha \\ \alpha & 1 \end{pmatrix}$.
As we show in Section~\ref{sec:gaussian_example} of the Appendix, the additive approximation has the
following closed form.
\begin{align*}
f^*_1(x_1) &= \left( \frac{T_1 - T_2 \alpha^2}{1 - \alpha^4} \right) x_1^2 + c_1\\
f^*_2(x_2) &= \left( \frac{T_2 - T_1 \alpha^2}{1 - \alpha^4} \right) x_2^2 + c_2
\end{align*}
Where $T_1 = H_{11} + 2H_{12} \alpha + H_{22} \alpha^2$, $T_2 = H_{22} + 2H_{12} \alpha + H_{11} \alpha^2$, $c_1, c_2$ are constants such that $f^*_1$ and $f^*_2$ both have mean zero. Let $H = \begin{pmatrix} 1.6 & 2 \\ 2 & 5\end{pmatrix}$, then it is easy to check that if $\alpha = - \frac{1}{2}$, then $f^*_1 = 0$
and additive faithfulness is violated, if $\alpha > \frac{1}{2}$, then $f^*_1$ is a concave function. We take the setting where $\alpha=-0.5$, compute the optimal additive functions via numerical simulation, and show the results in Figure~\ref{fig:gaussian_example}--$f^*_1$ is zero as expected.
\end{example}

Although the Gaussian distribution does not satisfy the boundary flatness condition, it is possible to approximate the Gaussian distribution arbitrarily well with distributions that do satisfy the boundary flatness conditions. We use the similar idea as that of Example~\ref{ex:mixture_approx}.

\begin{example} 
\label{examp:boundaryflat_example}
Let $\Sigma$ be as in Example~\ref{examp:gaussian_counterexample} with $\alpha = -0.5$ so that $f^*_1 = 0$. Consider a mixture $\lambda U[-(b+\epsilon), b+\epsilon]^2 + (1-\lambda) N_b(0, \Sigma)$ where $N_b(0,\Sigma)$ is the density of a \emph{truncated} bivariate Gaussian bounded in $[-b, b]^2$ and $U[-(b+\epsilon), b+\epsilon]^2$ is the uniform distribution over a square. The uniform distribution is supported over a slightly larger square to satisfy the boundary flatness conditions.

@@

\begin{figure*}[htp]
\vskip-10pt
	\centering
	\subfigure[Gaussian distribution]{
          \label{fig:gaussian_example}
		\centering
		{\includegraphics[width=0.4\textwidth]{figs/gaussian_example}}
	}
	\subfigure[Mixture approximation]{
          \label{fig:boundaryflat_example}
		\centering
		{\includegraphics[width=0.4\textwidth]{figs/boundaryflat_example}}
	}
\caption{Optimal additive projection of the quadratic function described in Example~\ref{examp:gaussian_counterexample} under both the Gaussian distribution described in Example~\ref{examp:gaussian_counterexample} and under the approximately Gaussian mixture distribution described in Example~\ref{examp:boundaryflat_example}. For the mixture approximation, we used $b=5, \epsilon=0.3, \lambda=0.0001$ where the parameters are defined in Example~\ref{examp:boundaryflat_example}. This example shows the effect and the importance of the boundary flatness conditions.}
\vskip-10pt
\end{figure*}

>>>>

The boundary flatness condition intuitively states that two
conditional densities $p(\mathbf{x}_{-j} \given x_j)$ and
$p(\mathbf{x}_{-j} \given x_j')$ are similar when $x_j$ and $x_j'$ are
both close to the same boundary point. It is thus much more general
than product densities. Boundary flatness is a weak condition because
it affects only an $\epsilon$-small region around the boundary;
$p(\mathbf{x}_{-j} \given x_j)$ can take arbitrary shapes away from
the boundary. Boundary flatness also allows arbitrary correlation
structure between the variables (provided $p(\mathbf{x}) >0$). In
Section~\ref{sec:boundary_flat}, we give a detailed discussion of the
boundary flatness condition and show examples of boundary flat
densities. In particular, we show that any density supported on a
compact set can be approximated arbitrarily well by boundary flat
densities.

@@

\begin{remark}
In Theorem~\ref{thm:convex_faithful}, we do not assume a parametric
form for the additive components; the additive approximations may not
be faithful if we take a parametric form. For example, suppose we
approximate a mean-zero convex function $f(X)$ by a linear form $X
\beta$. The optimal linear function in the population setting is
$\beta^* = \Sigma^{-1} \trm{Cov}(X, f(X))$ where $\Sigma$ is the
covariance matrix. Suppose the $X$s are independent, have a
symmetric distribution with unit variance, and suppose $f(\mathbf{x})
= x_1^2 - \E[X_1^2]$. Then $\beta^*_1 = \E[ X_1 f(X)] = \E[ X_1^3 -
  X_1 \E[X_1^2]] = 0$.
\end{remark}

@@

In this section, we give more examples of boundary flat densities (see
Definition~\ref{defn:boundary-point}) and discuss extending the notion
of boundary flatness to densities with a more general support. We
first start with a sufficient condition on the \emph{joint density}
that ensures boundary flatness.

@@

In the discussion so far, we have restricted our attention to
densities supported and positive on the hypercube $[0,1]^p$ to
minimize extraneous technical details. It may be possible to extend
the analysis to densities whose support is a convex and compact set so
long as the marginal density $p(x_j) > 0$ for all $x_j$ in the
support. A rigorous analysis of this however is beyond the scope of
this paper.

It may also possible to formulate a similar result to densities with
unbounded support, by using a limit condition $\lim_{|x_k| \rightarrow
  \infty} \frac{\partial p(\mathbf{\scriptstyle x}_{-k} \given
  x_k)}{\partial x_k} = 0$.  Such a limit condition however is not
obeyed by a correlated multivariate Gaussian distribution.  The next
example shows that certain convex functions are not additively
faithful under general multivariate Gaussian distributions.

\begin{example}
\label{examp:gaussian_counterexample}
Consider a two dimensional quadratic function $f( \mathbf{x}) = \mathbf{x}^\tran H \mathbf{x} + c$ with zero mean where $H = \begin{pmatrix} H_{11} & H_{12} \\ H_{12} & H_{22}\end{pmatrix}$ is positive definite and a Gaussian distribution $X \sim N(0, \Sigma)$ where $\Sigma = \begin{pmatrix}1 & \alpha \\ \alpha & 1 \end{pmatrix}$.
As we show in Section~\ref{sec:gaussian_example} of the Appendix, the additive approximation has the
following closed form.
\begin{align*}
f^*_1(x_1) &= \left( \frac{T_1 - T_2 \alpha^2}{1 - \alpha^4} \right) x_1^2 + c_1\\
f^*_2(x_2) &= \left( \frac{T_2 - T_1 \alpha^2}{1 - \alpha^4} \right) x_2^2 + c_2
\end{align*}
Where $T_1 = H_{11} + 2H_{12} \alpha + H_{22} \alpha^2$, $T_2 = H_{22}
+ 2H_{12} \alpha + H_{11} \alpha^2$, $c_1, c_2$ are constants such
that $f^*_1$ and $f^*_2$ both have mean zero. Let $H = \begin{pmatrix}
  1.6 & 2 \\ 2 & 5\end{pmatrix}$, then it is easy to check that if
  $\alpha = - \frac{1}{2}$, then $f^*_1 = 0$ and additive faithfulness
  is violated, if $\alpha > \frac{1}{2}$, then $f^*_1$ is a concave
  function. We take the setting where $\alpha=-0.5$, compute the
  optimal additive functions via numerical simulation, and show the
  results in Figure~\ref{fig:gaussian_example}.  Here $f^*_1$ is zero as
  expected.
\end{example}

Although the Gaussian distribution does not satisfy the boundary
flatness condition, it is possible to approximate the Gaussian
distribution arbitrarily well with distributions that do satisfy the
boundary flatness condition. We use an idea similar to that of
Example~\ref{ex:mixture_approx}.

\begin{example} 
\label{examp:boundaryflat_example}
Let $\Sigma$ be as in Example~\ref{examp:gaussian_counterexample} with
$\alpha = -0.5$ so that $f^*_1 = 0$. Consider a mixture $\lambda
U[-(b+\epsilon), b+\epsilon]^2 + (1-\lambda) N_b(0, \Sigma)$ where
$N_b(0,\Sigma)$ is the density of a \emph{truncated} bivariate
Gaussian bounded in $[-b, b]^2$ and $U[-(b+\epsilon), b+\epsilon]^2$
is the uniform distribution over a square. The uniform distribution is
supported over a slightly larger square to satisfy the boundary
flatness condition.

@@

\begin{figure*}[htp]
\vskip-10pt
	\centering
	\subfigure[Gaussian distribution]{
          \label{fig:gaussian_example}
		\centering
		{\includegraphics[width=0.4\textwidth]{figs/gaussian_example}}
	}
	\subfigure[Mixture approximation]{
          \label{fig:boundaryflat_example}
		\centering
		{\includegraphics[width=0.4\textwidth]{figs/boundaryflat_example}}
	}
\caption{Optimal additive projection of the quadratic function
  described in Example~\ref{examp:gaussian_counterexample} under both
  the Gaussian distribution described in
  Example~\ref{examp:gaussian_counterexample} and the
  approximately Gaussian mixture distribution described in
  Example~\ref{examp:boundaryflat_example}. For the mixture
  approximation, we used $b=5, \epsilon=0.3, \lambda=0.0001$ where the
  parameters are defined in
  Example~\ref{examp:boundaryflat_example}. This example shows the
  effect and the importance of the boundary flatness condition.}
\vskip-10pt
\end{figure*}
